---
title: "Independent Week 14"
author: "Christine Kandeo"
date: "2/5/2022"
output: html_document
---

 
# 1. Data Preparation

Loading Libraries
``` {r}
#Loading the relevant libraries:
library(tidyverse)
library(tibbletime)

#For anomaly detection:
library(anomalize)


```


Loading the Dataset
```{r}
#Importing the data to the Global Environment:
Sales <- read.csv("http://bit.ly/CarreFourSalesDataset", header = TRUE, sep = ",")

#Printing the first 4 rows of the dataframe
head(Sales, n=4)

```


```{r}
#Checking the Number of Rows and Columns:
dim(Sales)

```

- The dataset has 1,000 Rows and 2 Columns


Data Uniformity
```{r}
#Getting Information on the data types on each respective column:
sapply(Sales, class)

```

```{r}
#Converting the Date column from Character to Datetime:
Sales$Date <- as.Date(Sales$Date, "%m/%d/%Y")

```


```{r}
#Viewing the full information:
str(Sales)
```

Checking for null entries
```{r}
#Checking for null entries in each column:
colSums(is.na(Sales))

#For the entire data set: 
sum(is.na(Sales))

```
Duplicated entries
```{r}
#Checking for identical entries:
sum(duplicated(Sales))

```
The Data lacked Null or Duplicate entries


# Anomaly Detection

Involves considering the data points that are outliers or an exceptional event.


```{r}
#Detecting anomalies in the dataset:
Sales %>%
  group_by(Date) %>%
  summarise(Orders = sum(Sales)) %>%
  time_decompose(Orders, method = "twitter", frequency = "auto", trend = "auto") %>%
  anomalize(remainder, method = "gesd",alpha = 0.05, max_anoms = 0.2) %>%
  time_recompose() %>%
  plot_anomalies(
  time_recomposed =T,
  ncol = 3,
  color_no = "#2c3e50",
  color_yes = "#e31a1c",
  fill_ribbon = "grey70",
  size_dots = 2,
  size_circles = 4)
 
```


```{r}
#Visualizing the decomposition:
Sales %>%
  group_by(Date) %>%
  summarise(Orders = sum(Sales)) %>%
  time_decompose(Orders, method = "twitter", frequency = "auto", trend = "auto") %>%
  anomalize(remainder,method = "gesd", alpha = 0.05, max_anoms = 0.2) %>%
  time_recompose() %>%
  plot_anomaly_decomposition()

```

From the Plots above, the dataset has anomalies.

Outliers can be detected between February, March and April

