---
title: "Independent Week 14"
author: "Christine Kandeo"
date: "2/5/2022"
output: html_document
---

 
# 1. Data Preparation

Loading Libraries
``` {r}
#Loading the relevant libraries:
library(tidyverse)

#For association:
library(arules)

#For mapping rules:
library(arulesViz)


```


Loading the Dataset
```{r}
#Importing the data to the Global Environment:
Commodities <- read.transactions("http://bit.ly/SupermarketDatasetII", header = TRUE, sep = ",")

#Exploring the data using the Summary function:
summary(Commodities)

```
- The dataset has 119 columns with 7,500 rows/itemsets.

- The most frequent Item is Mineral Water.


```{r}
#Previewing the first five transactions:
inspect(Commodities[1:5])
```


```{r}
#Examine the frequency of the items:
itemFrequency(Commodities[, 1:3])

```

```{r}
#Plotting a FrequencyPlot to determine the frequency of items appearing in the Market Basket:
itemFrequencyPlot(Commodities, support = 0.025, cex.names=0.8, xlim = c(0,0.3),type = "relative", horiz = TRUE, col = "dark red", las = 1, topN = 10, 
xlab = paste("Proportion of Market Baskets Containing Item",
"\n(Item Relative Frequency or Support)"))

```

Mineral Water and Eggs are the top most purchased items in the store.


# Association Rule Mining

```{r}
#Using the default settings to see the number of rules learned:
apriori(Commodities)

```

Using the Default settings results in zero rules learned


```{r}
#Setting better support and confidence levels to learn more rules:
Commodityrules <- apriori(Commodities, parameter = list(support = 0.001, confidence = 0.7, 
    minlen = 2))

```

Using a Confidence of 0.7 and Support of 0.001 results in 198 Rules learned.


```{r}
#Evaluating the Model performance(summary of commodities association)
summary(Commodityrules)

```

```{r}
#Looking at the first three rules:
inspect(Commodityrules[1:3])

```
# Targeting Items


1. If customers buy Mineral Water, what are they likely to buy?

```{r}
#To find out which items are likely to be purchased with mineral water:
rules<-apriori(data=Commodities, parameter=list(supp=0.001,conf = 0.8), 
 appearance = list(default="lhs",rhs="mineral water"),
 control = list(verbose=F))
 rules<-sort(rules, decreasing=TRUE,by="confidence")
 inspect(rules[1:5])

 
```

Above are the products that are commonly bought with Mineral water.


# Assosiation Rule Visualization

Representing the association rules for commonly bought products with Mineral Water below in a network diagram:

```{r}
#A graph to map the rules:
plot(rules[1:5], method="graph")

```















