---
title: "Independent Week 14"
author: "Christine Kandeo"
date: "2/5/2022"
output: html_document
---

# DEFINING THE QUESTION

# Specific Data Analytics Question

Exploring and reducing the recent Carefour Marketing data using Principal Component Analysis and selecting the best features.


# Metrics for Success

Accurately determining the Principal Component's explaining most of the variability in the data and selecting the best features


# Understanding the Context

Carrefour Kenya are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest number of sales. This project aims to explore a recent marketing dataset by performing various unsupervised learning techniques and providing insights.

# Experimental Design.

1.Data Preparation

 -  Loading the dataset
 
 -  Data Uniformity
 
 -  Checking for Missing/Duplicate Values
 
 
2.Dimensionality Reduction

 - Principal Component Analysis
 
 
3.Feature Selection

 - Wrapper Method
 
 - Filter Method
 

4. Conclusions and Recommendations


  
# Data Relevance

 Invoice id - Computer generated sales slip invoice identification number

 Branch - Branch of the Supermarket (A, B and C)

 Customer type - Type of customers, recorded by Members for customers using member card and Normal for without member card

 Gender - Male or Female

 Product line - General item categorization groups: Electronic accessories, Fashion accessories, Food and beverages, Health and beauty, Home and lifestyle, Sports and travel

 Unit price - Price of each product

 Quantity - Number of products purchased by customer

 Tax - 5% tax fee for customer buying

 Date - Date of purchase (Record available from January 2019 to March 2019)

 Time - Purchase time

 Payment - Payment used by customer for purchase (3 methods are available – Cash, Credit card and Ewallet)

 COGS - Cost of goods sold

 Gross margin percentage - Gross margin percentage

 Gross income - Gross income

 Rating - Customer stratification rating on their overall shopping experience (On a scale of 1 to 10)
 
 Total - Total price including tax.
 
 
# 1. Data Preparation

Loading Libraries
``` {r}
#Loading the relevant libraries:
library(stats)
library(devtools)
library(ggbiplot)
library(caret)
library(GGally)

#For data manipulation:
library(dplyr)

#For visualization:
library(ggcorrplot)
library(ggplot2)

#For Feature selection:
library(clustvarsel)
library(mclust)
library(corrplot)

```


Loading the Dataset
```{r}
#Importing the data to the Global Environment:
Sales <- read.csv("http://bit.ly/CarreFourDataset", header = TRUE, sep = ",")

#Printing the first 4 rows of the dataframe
head(Sales, n=4)

```

```{r}
#Checking the Number of Rows and Columns:
dim(Sales)

```
The Dataset is made up of 1,000 Rows and 16 Columns


Data Uniformity
```{r}
#Getting Information on the data types on each respective column:
sapply(Sales, class)

```

```{r}
#Viewing the full information:
str(Sales)

```


Missing Values
```{r}
#Checking for null entries in each column:
colSums(is.na(Sales))

#For the entire data set: 
sum(is.na(Sales))

```
The dataset has no missing values


Duplicate values
```{r}
#Checking for identical entries:
sum(duplicated(Sales))

```
The dataset has no identical entries


# 2. Dimensionality Reduction

This process involves reducing the dimensions of the feature set. 


Technique: Principal Component Analysis

```{r}
#Checking PCA eligibility for linear dependency:
#Creating a dataframe Cr:
Dr<- Sales

#Visualizing the Plot
corr_map <- ggcorr(Dr[, 1:16], method=c("everything", "pearson"), label=TRUE, hjust = .90, size = 3, layout.exp = 2)
corr_map

```

The Plot above shows presence of Multicollinearity. PCA can effectively take advantage of Multicollinearity by combining the highly correlated variables into a set of uncorrelated variables.


```{r}
#Implementing PCA:
#Since PCA only accepts continuous variables, the categorical observations are ignored
#The Center argument is True so as to shift the variables to Zero centered
#Scale is true as the variables measurements are in different scales:

Pca <- prcomp(Sales[,c(6:8, 12, 14, 15)], center = TRUE, scale = TRUE)

#Displaying the information on Standard deviation and Loadings:
summary(Pca)

```
The First component explains 65% of the variability in the data set. The Second component explains 17% of total variability in the data set.

This indicates the first two components explain 82% of the Total variation in the data set.

```{r}
#Plotting a scree plot of the Eigenvalues:  
plot(Pca)
screeplot(Pca, type = "line", main = "Scree plot of Eigenvalues")

```

The Scree Plot returns a plot of the variances (y-axis) associated with the PC's (x-axis). 

It shows how many Pc's to retain for further analysis. In this case, 6 with the first 2 Pc's explaining most of the variability in the data.

```{r}
#Making a biplot, which includes both the position of each sample in terms of PC1 and PC2 and showing how the initial variables map onto this: 
ggbiplot(Pca, obs.scale = 0.8, var.scale = 1, groups = as.factor(Sales$Quantity), ellipse = TRUE, circle = TRUE)


```

The plot shows how samples relate to one another in the PCA based on the groups generated from quantity.


From the analysis, the first 2 Principal Components explain most of the variability in the data.


# 3. Feature Selection

This is the process of filtering irrelevant or redundant features from the dataset. 

Having irrelevant features in the data can decrease the accuracy of the models and make a model learn based on irrelevant features.

Techniques applied are:


1. Wrapper Method

```{r}
#Using the clustvarsel function to find the optimal subset of variables in the dataset: 
out = clustvarsel(Sales[,c(6:8, 12, 14, 15)])
out


```

The subset to be used for the Clustering Model is composed of Cogs and Unit Price


2. Filter Method

```{r}
#Using the findcorrelation function to create a subset of variables:
#Creating a dataframe Core:
Core <- subset(Sales, select = -c(Date, Time, Invoice.ID, Customer.type, Branch, Gender, Product.line, Payment, gross.margin.percentage, Total))


#Calculate correlation matrix:
correlationMatrix <- cor(Core)

#Finding attributes that are highly corrected (ideally >0.75):
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.75)

#Printing indexes of highly correlated attributes:
names(Core[,highlyCorrelated])


```
Tax and Cogs are the highly correlated features. They have a perfect positive correlation of 1.

```{r}
#Removing the Redundant Features:
Core2 <- Core[-highlyCorrelated]

#Visualizing the representation:
par(mfrow = c(1, 2))
corrplot(correlationMatrix, order = "hclust")
corrplot(cor(Core2), order = "hclust")

```

After removing redundant variables, the best Features are Rating, Unit Price, Quantity and Gross Income.

# 4. Conclusions and Recommendations

 - The filter measure is used as its proven to be fast to compute while still capturing the usefulness of the feature set.

 - The wrapper method on the other hand uses a predictive model to score feature subsets.Each new subset is used to train a model counting the number of mistakes made (Error rate of the model). This training makes the method Computationally Intensive but provides the best performing features.

 - The fact that the features selected from the Filter method are not tuned to a specific type of predictive model, The best recommended features are: Cost Of Goods Sold and Unit Price.

